---
title: "Model"
format: html
execute:
  echo: false
  warning: false
  message: false
---

## 📊 Logistic Regression Model

This section explains the logistic regression model used to predict **teen smartphone addiction** based on behavioral and psychological factors.

---

### 🧮 General Logistic Regression Equation

$$
\log\left( \frac{P(\text{Addicted} = 1)}{1 - P(\text{Addicted} = 1)} \right) =
\beta_0 + \beta_1 \cdot \text{DailyUsage} +
\beta_2 \cdot \text{SleepHours} +
\beta_3 \cdot \text{SocialMediaTime} +
\beta_4 \cdot \text{AnxietyLevel} +
\beta_5 \cdot \text{DepressionLevel}
$$

Where:

- **DailyUsage**: Daily smartphone usage, in hours  
- **SleepHours**: Average hours of sleep per night  
- **SocialMediaTime**: Hours spent on social media per day  
- **AnxietyLevel**: Self-reported anxiety score  
- **DepressionLevel**: Self-reported depression score  

---

### Estimated Logistic Regression Equation

The fitted model from your data yields the following equation:

$$
\log\left( \frac{P(\text{Addicted} = 1)}{1 - P(\text{Addicted} = 1)} \right) =
-2.869 + 1.111 \cdot \text{DailyUsage} -
0.580 \cdot \text{SleepHours} +
1.235 \cdot \text{SocialMediaTime} +
0.003 \cdot \text{AnxietyLevel} +
0.046 \cdot \text{DepressionLevel}
$$

This equation lets us estimate a teen's probability of addiction based on their habits and psychological state.

---

### 🔢 Model Code and Output Table

```{r}
# Load necessary libraries
library(readr)
library(dplyr)
library(parsnip)
library(broom)
library(knitr)

# Load the dataset
data <- read_csv("C:/Users/user/Desktop/projects/teen_phone_addiction_dataset.csv")

# Create a binary target variable (1 = addicted, 0 = not addicted)
data <- data |>
  mutate(addicted = if_else(Addiction_Level >= 8, 1, 0),
         addicted = factor(addicted))

# Fit logistic regression model
model <- logistic_reg() |>
  set_engine("glm") |>
  fit(addicted ~ Daily_Usage_Hours + Sleep_Hours + Time_on_Social_Media + Anxiety_Level + Depression_Level,
      data = data)

# Extract tidy summary of coefficients
tidy_model <- tidy(model$fit)

# Show clean regression table
kable(tidy_model, caption = "🧮 Logistic Regression Coefficients", digits = 3)
```

## ✅ Conclusion

The logistic regression table summarizes how each predictor influences the likelihood of teen smartphone addiction:

- **Intercept (-2.869):** Represents the baseline log-odds of being addicted when all predictors are zero.
- **Daily_Usage_Hours (1.111):** A strong positive coefficient indicates that each additional hour of daily phone use significantly increases the odds of addiction.
- **Sleep_Hours (-0.580):** The negative sign suggests that more sleep is associated with **lower odds** of addiction. For each extra hour of sleep, the log-odds of addiction decrease.
- **Time_on_Social_Media (1.235):** Like daily phone use, more time on social media is a strong predictor of addiction.
- **Anxiety_Level (0.003):** This coefficient is very small and not statistically significant (*p = 0.874*), meaning anxiety level doesn’t show a clear relationship with addiction in this model.
- **Depression_Level (0.046):** Though the effect is smaller, it is statistically significant (*p = 0.023*), suggesting that higher depression scores slightly increase the likelihood of addiction.

###  Key Takeaways:

- **Daily usage and social media time** are the most impactful predictors.
- **Sleep** plays a protective role against addiction.
- **Depression** matters, but **anxiety** does not appear significant in this sample.
- The p-values confirm that most predictors (except anxiety) are statistically meaningful at conventional thresholds (*p < 0.05*).

These insights can inform strategies for reducing phone addiction in teens by targeting usage habits, promoting better sleep, and supporting mental health—especially depression.


```{r}
# Load libraries
library(plotly)
library(dplyr)
library(readr)

# Read data
data <- read_csv("C:/Users/user/Desktop/projects/teen_phone_addiction_dataset.csv")

# Create binary addiction variable
data <- data |>
  mutate(addicted = if_else(Addiction_Level >= 8, 1, 0))

# Fit logistic regression model
model <- glm(
  addicted ~ Daily_Usage_Hours + Time_on_Social_Media + Sleep_Hours + Anxiety_Level + Depression_Level,
  data = data,
  family = "binomial"
)

# Generate grid for plotting
grid <- expand.grid(
  Daily_Usage_Hours = seq(min(data$Daily_Usage_Hours), max(data$Daily_Usage_Hours), length.out = 50),
  Time_on_Social_Media = seq(min(data$Time_on_Social_Media), max(data$Time_on_Social_Media), length.out = 50)
)

# Set median values for other predictors
grid$Sleep_Hours <- median(data$Sleep_Hours, na.rm = TRUE)
grid$Anxiety_Level <- median(data$Anxiety_Level, na.rm = TRUE)
grid$Depression_Level <- median(data$Depression_Level, na.rm = TRUE)

# Predict addiction probability
grid$prob <- predict(model, newdata = grid, type = "response")

# Reshape for 3D plot
z_matrix <- matrix(grid$prob, nrow = 50, byrow = TRUE)

# Create 3D surface plot
plot_ly(
  x = ~unique(grid$Daily_Usage_Hours),
  y = ~unique(grid$Time_on_Social_Media),
  z = ~z_matrix,
  type = "surface",
  colorscale = "Purples"
) |>
  layout(
    title = "3D Surface: Addiction Probability",
    scene = list(
      xaxis = list(title = "Daily Usage Hours"),
      yaxis = list(title = "Time on Social Media"),
      zaxis = list(title = "Probability of Addiction")
    )
  )

```